---
title: "Quarto example"
format: 
  html:
    toc: true
  pdf: default
editor: visual
date: last-modified # display the date of the last edit
author:
  - name: "M V"
    affiliation: "University of Bern"
    email: "m.v@unibe.ch"
  - name: "Co Author"
    affiliation: "University of Whatever"
execute:
  warning: false
---

## Data analysis

```{r}
#| warning: true
#| echo: true
library("dplyr") # data wrangling
library("ggplot2") # plotting
library("kableExtra") # tables
```

## Load the data from the statistics office of the city of Zurich and clean it

```{r}
url <- "https://data.stadt-zuerich.ch/dataset/bev_vornamen_baby_od3700/download/BEV370OD3700.csv"
dat <- read.csv(url)
datclean <- dat |>
    rename(year = StichtagDatJahr,
           name = Vorname,
           sex = SexLang,
           births = AnzGebuWir) |>
    mutate(sex = ifelse(sex == "weiblich", "female", "male"))
```

## most common name by year and sex

```{r}
summaries <- datclean |>
  select(year, name, sex, births) |>
  group_by(sex, year) |>
  slice_max(order_by = births, n = 1) |>
  arrange(-year)
```

See the most common names in @tbl-names.

```{r}
#| label: tbl-names
summaries |>
  kable(caption = "Most common names of newborns in the city of Zurich by year and sex")
```

## Plot the most common names between 2021 and 2025

```{r}
#| fig-cap: "Most common names of Zurich newborns in 2021-2025."
datclean |>
  # The upper bound is deliberate: the data set gains a year of records over
  # time, and bounding the window keeps this figure matching its caption.
  filter(year >= 2021, year <= 2025) |>
  group_by(sex, name) |>
  summarise(births = sum(births)) |>
  slice_max(order_by = births, n = 10) |>
  ggplot(aes(x = reorder(name, births), y = births)) +
  facet_wrap(~ sex, scales = "free") +
  geom_bar(stat = "identity") +
  labs(x = "Name", y = "Births", 
       title = "") +
  coord_flip() +
  theme_minimal()
```

## Frequency of names over time

```{r}
#| label: freq-over-time
datclean |>
  group_by(sex, name, year) |>
  filter(name %in% c("Fabio", "Daniel", "Michael")) |>
  ggplot(aes(x = year, y = births, color = name)) +
  geom_line() +
  geom_point() +
  scale_color_viridis_d() +
  labs(x = "Year", y = "Births", color = "") +
  expand_limits(y = 0) +
  theme_minimal()
```

## Example for inline code

In 2023, the most common male name was `r filter(summaries, year == 2023, sex == "male")$name` whereas the most common female name was `r filter(summaries, year == 2023, sex == "female")$name`.

Problem with ties: In 2025, the most common male name was `r filter(summaries, year == 2025, sex == "male")$name` whereas the most common female name was `r filter(summaries, year == 2025, sex == "female")$name`.

## R session info

```{r}
#| label: sessioninfo
sessionInfo()
```
